Behind Bars: Statistical Insights into the Wellbeing of U.S. Prisoners

### 2024-02-16 paper/paper.qmd
added the updated abstract

---
title: "Behind Bars: Statistical Insights into the Wellbeing of U.S. Prisoners"
author:
  - Monica Sainani
  - Kaavya Kalani
  - Aamishi Avarsekar
thanks: "Code and data are available at: https://github.com/kaavyakalani26/prison_health.git and replication aspects are available at: https://www.socialsciencereproduction.org/reproductions/82b709e2-4aa0-40e8-8abe-195c51ae739b/index"
date: today
date-format: long
abstract: "This research endeavors to scrutinize the well-being of individuals within the prison system, focusing on the prevalence of physical diseases, mental health issues, and mortality rates among incarcerated populations. The prison environment presents unique challenges that can significantly impact the overall health of individuals, making it imperative to investigate the multifaceted dimensions of prisoners' well-being. Utilizing data collected over the past 15 years, this study employs a comprehensive analysis to discern patterns, trends, and potential variations in the health outcomes of prisoners. By examining the intersection of physical and mental health alongside mortality rates, the research aims to contribute valuable insights into the current state of prisoners' well-being. This analysis not only serves to inform our understanding of the challenges faced by incarcerated individuals but also provides a foundation for evidence-based policy recommendations aimed at fostering improved conditions within correctional facilities. Through this exploration, we strive to contribute to the ongoing dialogue surrounding the intersection of criminal justice and public health, emphasizing the importance of compassionate and effective strategies for promoting the well-being of those within the prison system."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

>>>>

---
title: "Behind Bars: Statistical Insights into the Wellbeing of U.S. Prisoners"
author:
  - Monica Sainani
  - Kaavya Kalani
  - Aamishi Avarsekar
thanks: "Code and data are available at: https://github.com/kaavyakalani26/prison_health.git and replication aspects are available at: https://www.socialsciencereproduction.org/reproductions/82b709e2-4aa0-40e8-8abe-195c51ae739b/index"
date: today
date-format: long
abstract: "This research explores the causes, consequences, and demographics of prisoner deaths in the United States.
It reveals an increasing trend in the death of prisoners, highlighting heart problems and cancer as major causes.
The frequency of cancer is shown by demographic study, with Black, Hispanic, and White inmates being most affected.
The significant impact that prison staff has on the wellbeing of inmates is highlighted by gender-based differences
in the availability of medical treatment and institutional discrimination.

The study acknowledges the shortcomings in demographic representation and argues for a larger dataset.
The study intends to shed light on aspects such as racism and hate crimes in correctional settings, underscoring the
need for inclusive statistics on racial and sexual minorities. The request for more datasets is meant to strengthen the
 research and provide a more comprehensive comprehension of the complex relationship between conditions in jail and overall health.

In addition, the report recommends cross-national comparisons and an analysis of the criminal justice systems in
countries with similar populations in order to evaluate general well-being and pandemic readiness. The study provides
insight into the difficulties faced by prisoners and establishes a foundation for further investigations into systemic
problems and healthcare disparities in correctional environments."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---
